The first isomorphism theorem is one of the classical results in universal algebra.
While this statement is usually about algebraic structures in sets \cite{lynge2019}, we look at a generalization to the 1-truncated case.
More specifically, we formulate and prove this theorem for algebras for the signatures defined in Definition \ref{def:signature}.

Before stating and proving the isomorphism theorem, we need to generalize several notions from universal algebra to the 1-truncated case.
First of all, following the approach of Section \ref{sec:finite_limits}, we define the image of an algebra morphism using displayed algebras.
Second of all, we define congruence relations and their quotients.
The main idea here is to use Construction \ref{constr:alg_biadj}, where
we lifted the groupoid quotient to a biadjunction from $\algG{\sign}$ to $\algebra{\sign}$.
This result also indicates how congruence relations will be defined.
Basically, a congruence relation on $X : \algebra{\sign}$ is a groupoid structure on $X$ which gives an algebra of groupoids.
Once these notions are in place, we can formulate and prove the first isomorphism theorem similarly to the set-theoretical version.

\subsection{The Image}
We start by defining the image of an algebra morphism.
For each $f : A \onecell B$ we are after a factorization $A \onecell \Im{f} \onecell B$.

\begin{problem}
\label{prob:image}
Given algebras $A$ and $B$ for a signature $\sign$ and an algebra morphism $f : A \onecell B$,
to construct an algebra $\Im(f)$ and morphisms $\ImProj{f} : A \onecell \Im{f}$ and $\ImInc{f} : \Im(f) \onecell B$. 
\end{problem}

\begin{construction}{prob:image}
We define $\Im(f)$ to be the total algebra $\TotalAlg Y$ of the following displayed algebra $Y$ over $B$:
\begin{itemize}
\item The underlying family of 1-types over $B$ is $Y(x) \eqdef \exists a : A.\ f(a) = x$.
\item For each $x : \polyAct{\pointconstr}{A}$, we are required to construct
a map $\DispAlgPoint{Y} : \polyDact{\pointconstr}{Y}(x) \rightarrow
\exists a : A.\ f(a) = \AlgPoint{B}(x)$. Suppose we have $\pover{x}
: \polyDact{\pointconstr}{Y}(x)$. By induction on the polynomial
$\pointconstr$, it is possible to derive from $\pover{x}$ an
inhabitant of $\exists y
: \polyAct{\pointconstr}{A}.\ \polyAct{\pointconstr}{f}(y) = x$. By
invoking the elimination principle of propositional truncation, it is sufficient to
construct a map $f : \Sum {y
: \polyAct{\pointconstr}{A}}.\ \polyAct{\pointconstr}{f}(y) =
x \rightarrow \Sum {a : A}.\ f(a) = \AlgPoint{B}(x)$ in
order to define the map $\DispAlgPoint{Y}$.  So, assume that we have
$y : \polyAct{\pointconstr}{A}$ and a path $p
: \polyAct{\pointconstr}{f}(y) = x$. Take $f(y,p)$ to be the pair
consisting of $\AlgPoint{A}(y)$ and the path
\[
f(\AlgPoint{A}(y))
\stackrel{\AlgMapPoint{f}(y)}{=}
\AlgPoint{B}(\polyAct{\pointconstr}{f}(y))
\stackrel{\ap {\AlgPoint{B}} p}{=}
\AlgPoint{B}(x)
\]
\item Since $Y(x)$ is a proposition for all $x : B$, the construction of
paths $\DispAlgPath{Y}{j}$ and globes $\DispAlgHomot{Y}{j}$ is
straightforward.
\end{itemize}

The map $\ImInc{f}$ is the first projection and $\ImProj{f}$ sends $x : A$ to
$(f(x),\PC(x,\refl{f(x)}))$, where $\PC$ is the point constructor of
propositional truncation. 
\end{construction}

\subsection{Congruence Relations}
Next we define congruence relations and we show that each such relation gives an algebra in groupoids.
The difficulty here is constructing the homotopy constructor of that algebra.
For this reason, we define congruence relations in two steps, and we start by defining \emph{path congruence relations} and showing that these give rise to path algebras in groupoids.

\begin{definition}
Let $\sign$ be a signature and let $X : \algebra{\sign}$.
A \fat{path congruence relation} $R$ on $X$ consists of
\begin{itemize}
	\item a groupoid structure $R$ on $X$;
	\item for each $x, y : \polyAct{\pointconstr[\sign]}{X}$ and $f : \mor{\polyAct{\pointconstr[\sign]}{R}}{x}{y}$, a morphism $\AlgPoint{R} : \mor{R}{\AlgPoint{X}(x)}{\AlgPoint{X}(y)}$ and a proof that this assignment is functorial;
	\item a proof that the assignment $\Lam x, \idtoiso(\AlgPath{X}{j}(x))$ gives a natural transformation $\semEG{\pathleft[\sign]_j} \twocell \semEG{\pathright[\sign]_j}$ for each $j : \pathlabel[\sign]$.
\end{itemize}
\end{definition}

\begin{problem}
\label{prob:path_congruence_relation_to_grpd}
Given a path congruence relation $R$,
to construct $\toGrpdPathAlg{R} : \pathalgG{\sign}$.
\end{problem}

\begin{construction}{prob:path_congruence_relation_to_grpd}
We only discuss the data involved.
\begin{itemize}
	\item The carrier $\toGrpdPathAlg{R}$ is the groupoid whose type of objects is $X$ and whose morphisms are given by $R$.
	\item Next we define a functor $\semPG{\pointconstr[\sign]}(\toGrpdPathAlg{R}) \onecell \toGrpdPathAlg{R}$. It is defined to be $\AlgPoint{X}$ on objects and $\AlgPoint{R}$ on morphisms.
	\item Lastly, we define a natural transformation $\semEG{\pathleft[\sign]_j} \twocell \semEG{\pathright[\sign]_j}$ for each $j : \pathlabel[\sign]$. The component function of this transformation is $\Lam x, \idtoiso(\AlgPath{X}{j}(x))$. \qedhere
\end{itemize}
\end{construction}

To show that $\toGrpdPathAlg{R}$ is an algebra, we also need to give the homotopy constructor.
Since this constructor has both a point and a path constructor, we need to check an equality of morphisms of $\toGrpdPathAlg{R}$, which depend on points of $X$ and morphisms of $\toGrpdPathAlg{R}$.
As a result, we cannot reuse the homotopy constructor of $X$, because it only depends on points of $X$ and paths in $X$.
This means that to construct the groupoid algebra, the homotopy constructor needs to be constructed from scratch.
This is reflected in the following definition.

\begin{definition}
A path congruence relation $R$ is a \fat{congruence} relation if $\toGrpdPathAlg{R}$ is an algebra in groupoids.
\end{definition}

Note that each congruence relation $R$ gives rise to an algebra $\toGrpdAlg{R} : \algG{\sign}$.
By Construction \ref{constr:alg_biadj}, the groupoid quotient lifts to $\alggquot : \pseudo(\algG{\sign}, \algebra{\sign})$.
Hence, we can construct an algebra $\alggquot(\toGrpdAlg{R}) : \algebra{\sign}$ from a congruence relation $R$.
Note that the map $\gcl : X \rightarrow \alggquot(\toGrpdAlg{R})$ is a homomorphism of algebras.
Since $\alggquot$ is a left biadjoint, we also get a mapping property for $\alggquot(\toGrpdAlg{R})$.

\begin{remark}
\label{remark:mapping_property_congruence}
Suppose that we have two algebras $X, Y : \algebra{\sign}$ and a congruence relation $R$ on $X$.
From the biadjunction, we get a unique morphism $\overline{g} : \alggquot{\toGrpdAlg{R}} \onecell Y$ from a functor $g : R \onecell \algpgrpd{Y}$ such that $\overline{g}(\gcl(x)) = g(x)$ and $\ap{\overline{g}}{(\gcleq(p))} = g(p)$ where $p$ is a morphism in $R$ from $x$ to $y$.

Using this, we can give conditions for when an algebra morphism $f : X \onecell Y$ factors through $\alggquot{\toGrpdAlg{R}}$.
We can unfold the definition to find the necessary ingredients for the factorization.
For example, we need to show that $f$ lifts to a functor, which means we need to provide a path $f_2 : f(x) = f(y)$ for each $x, y : X$ and $p : \mor{R}{x}{y}$, and equalities $f_2(\idpath(x)) =\idpath(f_2(x))$ and $f_2(p \cdot q) = f_2(p) \vcomp f_2(q)$.
\end{remark}

\subsection{The First Isomorphism Theorem}
Now let us prove a generalization of the first isomorphism theorem to the 1-truncated case.
Note that our proof follows the same steps as the proof of the first isomorphism theorem for sets.
Let us start by characterizing adjoint equivalences in $\algM(\sign)$.

\begin{proposition}
\label{prop:algebra_adjequiv}
Suppose we have a signature $\sign$, algebras $X, Y : \algM(\sign)$, and an algebra morphism $f : X \onecell Y$.
Then $f$ is an adjoint equivalence in $\algM(\sign)$ if its carrier is an adjoint equivalence in $\onetypes$.
\end{proposition}

Before we state and prove the isomorphism theorem, we define the kernel of algebra morphisms.

\begin{definition}[Kernel]
Given a signature $\sign$, algebras $X, Y : \algM(\sign)$, and an algebra morphism $f : X \onecell Y$, we define a groupoid algebra $\Ker(f)$ on $X$, called the \fat{kernel} of $f$, by setting $R(x, y) \eqdef f(x) = f(y)$.
\end{definition}

\begin{theorem}\label{thm:iso_thm}
Let $\sign$ be a signature and suppose we have algebras $X, Y : \algM(\sign)$ and an algebra morphism $f : X \onecell Y$.
Then
\begin{itemize}
	\item there is a unique map $\overline{f} : \alggquot(\toGrpdAlg{\Ker(f)}) \onecell Y$ such that we have propositional equalities $p : \overline{f}(\gcl(x)) = f(x)$ for all $x$ and $\concat{\ap{\overline{f}}{(\gcleq(r))}}{p(y)} = \concat{p(x)}{r}$ where $r : f(x) = f(y)$;
	\item there is an adjoint equivalence $\widetilde{f} : \alggquot(\toGrpdAlg{\Ker(f)}) \onecell \Im(f)$ such that $\pi_1(\widetilde{f}(x)) = \overline{f}(x)$.
\end{itemize}
\end{theorem}

\begin{proof}
Using Remark \ref{remark:mapping_property_congruence}, we can define the maps $\overline{f} : \alggquot(\toGrpdAlg{\Ker(f)}) \onecell Y$ and $\widetilde{f} : \alggquot(\toGrpdAlg{\Ker(f)}) \onecell \Im(f)$.
Note that from Remark \ref{remark:mapping_property_congruence}, we also get that $\overline{f}$ satisfies the required equalities and that $\overline{f}$ is unique.
To show that $\overline{f}$ is an adjoint equivalence, we use Proposition \ref{prop:algebra_adjequiv}, and we do that by showing that the fibers are contractible.
Proving that the fibers are inhabited and propositional is similar to proving the surjectivity and injectivity for the set-theoretical first isomorphism theorem \cite{lynge2019}.
\end{proof}